---
title: "Helping and hindering as counterfactual difference-making"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: cosmo
    highlight: tango
---

# Setup

```{r setup, include = F, message = F, warning = F}
require(knitr)         # for knitting
library(ggstatsplot)   # extension of ggplot2
library(lme4)          # for mixed effects models
library(lmerTest)      # for p values in lmer
library(DT)            # for datatable() function
library(brms)          # for Bayesian regression models
library(Metrics)       # for RMSE
library(Hmisc)         # for bootstrapped confidence intervals
library(xtable)        # for converting to LaTex tables
library(scales)        # hue palette
library(showtext)      # for fonts
library(broom.mixed)   # for augmenting fitted models
library(ggrepel)       # avoid overlap in geom_text
library(patchwork)     # for multiple plots
library(png)           # for working with images
library(grid)          # for grob
library(egg)           # for geom_custom
library(ggtext)        # for styling ggplot text
library(ggnewscale)    # for multiple scales in the same plot
library(tidyverse)     # for everything else

knitr::opts_chunk$set(echo = T, warning = F, message = F)

# set default plot theme
theme_set(theme_classic() +
            theme(text = element_text(size = 14)))

# suppress warnings about grouping
options(dplyr.summarise.inform = F)
options(xtable.floating = F)
options(xtable.timestamp = "")
options(readr.show_col_types = F)

set.seed(1)

data_dir    = '../data/'
figures_dir = '../figures/'
info_dir    = function(exp) str_c('../model/trial_info/exp', exp, '/')
```

# Helper functions

## Data processing

```{r}
read_trials = function(exp, condition) {
  read_csv(str_c(data_dir, 'experiment', exp, '/', condition, '/trials.csv')) %>%
    filter(trial > 0) %>%
    drop_na(trial) %>%
    mutate(id = factor(id))
}

read_participants = function(exp, condition) {
  read_csv(str_c(data_dir, 'experiment', exp, '/', condition, '/participants.csv')) %>%
    mutate(id = factor(id))
}
```

## Demographics

```{r}
print_demographics = function(data) {
  data %>% group_by(gender) %>% summarise(n = n()) %>% print()
  print(sprintf('age: M = %.0f, SD = %.0f',
                mean(data$age, na.rm = T),
                sd(data$age, na.rm = T)))
  data %>% group_by(race) %>% summarise(n = n()) %>% print()
  data %>% group_by(ethnicity) %>% summarise(n = n()) %>% print()
  print(sprintf('time: %.1f minutes (SD = %.1f)',
                mean(data$time, na.rm = T)/60,
                sd(data$time, na.rm = T)/60))
}
```

## Scatter plot

```{r}
scatterplot = function(data, x, x_label, y, y_label) {
  r    = cor(data[[str_c(x, "_mean")]], data[[str_c(y, "_mean")]])
  rmse = rmse(data[[str_c(x, "_mean")]], data[[str_c(y, "_mean")]])

  g = ggplot(data = data,
             aes(x = get(str_c(x, "_mean")),
                 y = get(str_c(y, "_mean")))) +
    geom_abline(intercept = 0, slope = 1, linetype = 2, size = 0.2) +
    geom_linerange(size = 0.7,
                   mapping = aes(ymin = get(str_c(y, "_low")),
                                 ymax = get(str_c(y, "_high"))),
                   color = 'lightgray') +
    geom_errorbarh(mapping = aes(xmin = get(str_c(x, "_low")),
                                 xmax = get(str_c(x, "_high"))),
                   color = 'lightgray', height = 0) +
    geom_point(size = 1.5, fill = 'black') +
    geom_text(label = sprintf('RMSE = %.2f\nr = %.2f', rmse, r),
              hjust = 0, x = 0, y = 95, size = 4,
              check_overlap = T) +
    scale_x_continuous(name = x_label, limits = c(0, 100)) +
    scale_y_continuous(name = y_label, limits = c(0, 100)) +
    coord_fixed(ratio = 1, clip = 'off') +
    theme(legend.position = "none",
          text = element_text(size = 16),
          axis.title.x = element_markdown(size = 16),
          axis.title.y = element_markdown(size = 16),
          aspect.ratio = 1)

  return(g)
}
```

# Experiment 1

## Data

### Trial info

```{r}
trial.info.exp1 = read_csv(str_c(info_dir(1), 'model_predictions.csv')) %>%
  rename(cf_model = counterfactual,
         int_model = intention,
         effort_model = effort) %>%
  mutate(outcome = factor(outcome, levels = c('success', 'failure')),
         cf_model = cf_model * 100,
         int_model = int_model * 100,
         effort_model = 10*effort_model,
         cf_model_recode = ifelse(outcome == 'success',
                                  100 - cf_model, cf_model),
         int_model_recode = ifelse(outcome == 'success',
                                   int_model, 100 - int_model))

images.exp1 = data.frame(trial = 1:nrow(trial.info.exp1)) %>%
  mutate(grob = map(
    .x = trial,
    .f = ~ readPNG(str_c(figures_dir, 'trial_stills/experiment1/', .x, '.png'))))
```

### Responsibility

Participants were asked: “What factors influenced how you decided to respond? Do you have any questions or comments regarding the experiment?”
```{r}
data.exp1.resp = read_trials(1, 'responsibility')
data.exp1.resp.participants = read_participants(1, 'responsibility')

data.exp1.resp.participants %>%
  select(id, feedback) %>%
  datatable(rownames = F)
```

### Counterfactual

```{r}
data.exp1.cf = read_trials(1, 'counterfactual') %>%
  left_join(trial.info.exp1 %>%
              select(trial, outcome),
            by = 'trial') %>%
  mutate(cf_recode = ifelse(outcome == 'success',
                            100 - response, response))

data.exp1.cf.participants = read_participants(1, 'counterfactual')

data.exp1.cf.participants %>%
  select(id, feedback) %>%
  datatable(rownames = F)
```

### Intention

```{r}
data.exp1.int = read_trials(1, 'intention') %>%
  left_join(trial.info.exp1 %>%
              select(trial, outcome),
            by = 'trial') %>%
  mutate(int_recode = ifelse(outcome == 'success',
                             response, 100 - response))

data.exp1.int.participants = read_participants(1, 'intention')

data.exp1.int.participants %>%
  select(id, feedback) %>%
  datatable(rownames = F)
```

### Effort

```{r}
data.exp1.eff = read_trials(1, 'effort')
data.exp1.eff.participants = read_participants(1, 'effort')

data.exp1.eff.participants %>%
  select(id, feedback) %>%
  datatable(rownames = F)
```

## Demographics

```{r}
print_demographics(data.exp1.resp.participants %>%
                     rbind(data.exp1.cf.participants) %>%
                     rbind(data.exp1.int.participants) %>%
                     rbind(data.exp1.eff.participants))
```

## Analysis

### Model components

```{r}
list(cf  = list(judgment = data.exp1.cf %>%
                  group_by(trial) %>%
                  summarise(mean = mean(response)),
                model = trial.info.exp1 %>%
                  select(trial, mean = cf_model)),
     int = list(judgment = data.exp1.int %>%
                 group_by(trial) %>%
                 summarise(mean = mean(response)),
               model = trial.info.exp1 %>%
                 select(trial, mean = int_model)),
    eff = list(judgment = data.exp1.eff %>%
                 group_by(trial) %>%
                 summarise(mean = mean(response)),
               model = trial.info.exp1 %>%
                 select(trial, mean = effort_model))) %>%
  map_dfr(.id = 'condition',
          function(x) {
            d = left_join(x$judgment, x$model,
                          by = 'trial',
                          suffix = c('_judgment', '_model'))
            tibble(r = cor(d$mean_judgment, d$mean_model),
                   rmse = rmse(d$mean_judgment, d$mean_model))}
          )
```

### Responsibility model

Setting up data:
```{r}
data.exp1.modeling = data.exp1.resp %>%
  left_join(data.exp1.cf %>%
              group_by(trial) %>%
              summarise(cf_recode = mean(cf_recode)),
            by = 'trial') %>%
  left_join(data.exp1.int %>%
              group_by(trial) %>%
              summarise(int_recode = mean(int_recode)),
            by = 'trial') %>%
  left_join(data.exp1.eff %>%
              group_by(trial) %>%
              summarise(effort = mean(response)),
            by = 'trial') %>%
  left_join(trial.info.exp1, by = 'trial')
```

Model fitting:
```{r}
fit.exp1.cf = brm(
  formula = resp ~ 1 + cf_recode + (1 | id),
  data = data.exp1.modeling,
  iter = 2000,
  seed = 1,
  file = "cache/fit.exp1.cf")

summary(fit.exp1.cf)

fit.exp1.int = brm(
  formula = resp ~ 1 + int_recode + (1 | id),
  data = data.exp1.modeling,
  iter = 2000,
  seed = 1,
  file = "cache/fit.exp1.int")

summary(fit.exp1.int)

fit.exp1.cf_int = brm(
  formula = resp ~ 1 + cf_recode + int_recode + (1 | id),
  data = data.exp1.modeling,
  iter = 2000,
  seed = 1,
  file = "cache/fit.exp1.cf_int")

summary(fit.exp1.cf_int)

fit.exp1.cf_eff = brm(
  formula = resp ~ 1 + cf_recode + effort + (1 | id),
  data = data.exp1.modeling,
  iter = 2000,
  seed = 1,
  file = "cache/fit.exp1.cf_eff")

summary(fit.exp1.cf_eff)

fit.exp1.heuristic = brm(
  formula = resp ~ 1 + outcome + red_steps + blue_steps + box_steps + (1 | id),
  data = data.exp1.modeling,
  iter = 2000, seed = 1,
  file = "cache/fit.exp1.heuristic")

summary(fit.exp1.heuristic)
```

Model comparison:
```{r}
fit.exp1.cf = add_criterion(
  fit.exp1.cf,
  criterion = "loo",
  reloo = T,
  file = "cache/fit.exp1.cf")

fit.exp1.int = add_criterion(
  fit.exp1.int,
  criterion = "loo",
  reloo = T,
  file = "cache/fit.exp1.int")

fit.exp1.cf_int = add_criterion(
  fit.exp1.cf_int,
  criterion = "loo",
  reloo = T,
  file = "cache/fit.exp1.cf_int")

fit.exp1.cf_eff = add_criterion(
  fit.exp1.cf_eff,
  criterion = "loo",
  reloo = T,
  file = "cache/fit.exp1.cf_eff")

fit.exp1.heuristic = add_criterion(
  fit.exp1.heuristic,
  criterion = "loo",
  reloo = T,
  file = "cache/fit.exp1.heuristic")

loo_compare(fit.exp1.cf,
            fit.exp1.int,
            fit.exp1.cf_int,
            fit.exp1.cf_eff,
            fit.exp1.heuristic)
```

Individual model fitting and comparison:
```{r eval = F}
fit.exp1.cf_individual = brm(
  formula = resp ~ 1 + cf_recode,
  data = data.exp1.modeling %>% filter(id == 1),
  seed = 1,
  file = "cache/fit.exp1.cf_individual")

fit.exp1.int_individual = brm(
  formula = resp ~ 1 + int_recode,
  data = data.exp1.modeling %>% filter(id == 1),
  seed = 1,
  file = "cache/fit.exp1.int_individual")

fit.exp1.cf_int_individual = brm(
  formula = resp ~ 1 + cf_recode + int_recode,
  data = data.exp1.modeling %>% filter(id == 1),
  seed = 1,
  file = "cache/fit.exp1.cf_int_individual")

fit.exp1.cf_effort_individual = brm(
  formula = resp ~ 1 + cf_recode + effort,
  data = data.exp1.modeling %>% filter(id == 1),
  seed = 1,
  file = "cache/fit.exp1.cf_effort_individual")

fit.exp1.heuristic_individual = brm(
  formula = resp ~ 1 + outcome + red_steps + blue_steps + box_steps,
  data = data.exp1.modeling %>% filter(id == 1),
  seed = 1,
  file = "cache/fit.exp1.heuristic_individual")

data.exp1.fits = data.exp1.modeling %>%
  group_by(id) %>% nest() %>% ungroup() %>%
  mutate(
    fit_cf = map(.x = data,
                 .f = ~ update(fit.exp1.cf_individual,
                               newdata = .x,
                               seed = 1)),
    fit_int = map(.x = data,
                  .f = ~ update(fit.exp1.int_individual,
                                newdata = .x,
                                seed = 1)),
    fit_cf_int = map(.x = data,
                     .f = ~ update(fit.exp1.cf_int_individual,
                                   newdata = .x,
                                   seed = 1)),
    fit_cf_effort = map(.x = data,
                        .f = ~ update(fit.exp1.cf_effort_individual,
                                      newdata = .x,
                                      seed = 1)),
    fit_heuristic = map(.x = data,
                        .f = ~ update(fit.exp1.heuristic_individual,
                                      newdata = .x,
                                      seed = 1))) %>%
  mutate(
    fit_cf = map(.x = fit_cf,
                 .f = ~ add_criterion(.x, criterion = "loo")),
    fit_int = map(.x = fit_int,
                  .f = ~ add_criterion(.x, criterion = "loo")),
    fit_cf_int = map(.x = fit_cf_int,
                     .f = ~ add_criterion(.x, criterion = "loo")),
    fit_cf_effort = map(.x = fit_cf_effort,
                        .f = ~ add_criterion(.x, criterion = "loo")),
    fit_heuristic = map(.x = fit_heuristic,
                        .f = ~ add_criterion(.x, criterion = "loo")),
    model_comparison = pmap(.l = list(cf = fit_cf,
                                      int = fit_int,
                                      cf_int = fit_cf_int,
                                      cf_effort = fit_cf_effort,
                                      heuristic = fit_heuristic),
                            .f = ~ loo_compare(..1, ..2, ..3, ..4, ..5)),
    best_model = map_chr(.x = model_comparison,
                         .f = ~ rownames(.) %>% .[1]),
    best_model = factor(best_model,
                        levels = c("..1", "..2", "..3", "..4", "..5"),
                        labels = c("cf", "int", "cf_int", "cf_effort", "heuristic")))

save(list = c("data.exp1.fits"),
     file = 'cache/exp1.model_fits.RData')
```

```{r}
load(file = 'cache/exp1.model_fits.RData')

data.exp1.fits %>%
  count(best_model) %>%
  arrange(desc(n))
```

## Figures

```{r}
exp1.conditions = c('int', 'cf', 'eff', 'resp')

exp1.condition_labels = c('int' = 'Intention',
                          'cf' = 'Counterfactual',
                          'eff' = 'Effort',
                          'resp' = 'Responsibility')

exp1.condition_colors = c('int' = 'tan2',
                          'cf' = 'orchid3',
                          'eff' = 'aquamarine3',
                          'resp' = 'deepskyblue')

data.exp1.long = data.exp1.resp %>%
  mutate(condition = 'resp') %>%
  rbind(data.exp1.int %>%
          select(-outcome, -int_recode) %>%
          mutate(condition = 'int')) %>%
  rbind(data.exp1.cf %>%
          select(-outcome, -cf_recode) %>%
          mutate(condition = 'cf')) %>%
  rbind(data.exp1.eff %>%
          mutate(condition = 'eff')) %>%
  mutate(condition = factor(condition, levels = exp1.conditions))

data.exp1.model = trial.info.exp1 %>%
  select(trial, int = int_model,
         cf = cf_model, eff = effort_model) %>%
  left_join(data.exp1.modeling %>%
              cbind(fit.exp1.cf_int %>%
                      fitted(newdata = data.exp1.modeling,
                             re_formula = NA) %>%
                      as_tibble() %>%
                      select(pred = Estimate)) %>%
              distinct(trial, pred) %>%
              rename(resp = pred),
            by = 'trial') %>%
  pivot_longer(cols = -trial,
               names_to = 'condition',
               values_to = 'prediction') %>%
  mutate(condition = factor(condition, levels = exp1.conditions))
```

### All trials

```{r fig.width = 12, fig.height = 16}
ggplot(data = data.exp1.long,
       mapping = aes(x = condition,
                     y = response,
                     fill = condition)) +
  # mean judgments
  stat_summary(fun = "mean",
               geom = "bar",
               color = "black",
               alpha = 0.8,
               linewidth = 0.1) +
  # individual participant judgments
  geom_point(size = 1,
             stroke = 0.5,
             shape = 21,
             color = 'gray40',
             alpha = 0.5,
             position = position_jitter(height = 0, width = 0.2),
             show.legend = F) +
  # error bars
  stat_summary(fun.data = mean_cl_boot,
               geom = "linerange",
               color = "black",
               linewidth = 1) +
  # model predictions
  geom_point(data = data.exp1.model,
             mapping = aes(y = prediction),
             size = 3,
             stroke = 0.6,
             shape = 21,
             show.legend = F) +
  # trial images
  geom_custom(data = images.exp1,
              inherit.aes = F,
              mapping = aes(data = grob,
                            x = -Inf, y = Inf),
              grob_fun = function(x) rasterGrob(
                x, interpolate = T,
                vjust = 0, hjust = 0)) +
  # trial image numbers
  geom_text(aes(label = trial, x = 0.75, y = 124),
            color = 'black', size = 6,
            check_overlap = T) +
  facet_wrap(~ trial, nrow = 5, scales = 'free_x') +
  scale_fill_manual(values = exp1.condition_colors,
                    labels = exp1.condition_labels) +
  scale_y_continuous(name = 'Judgments',
                     breaks = seq(0, 100, 25),
                     expand = expansion(mult = 0.01)) +
  coord_cartesian(clip = "off",
                  ylim = c(0, 110)) +
  theme(text = element_text(size = 16),
        legend.position = 'bottom',
        legend.direction = 'horizontal',
        legend.text = element_text(
          size = 16,
          margin = margin(r = 1, l = 0.2, unit = 'cm')),
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        panel.grid.major.y = element_line(),
        panel.spacing.y = unit(4.1, "cm"),
        plot.margin = margin(t = 3.8, l = 0, r = 0, b = 0, unit = 'cm'))

# ggsave(str_c(figures_dir, 'exp1_bar_all.pdf'),
#        width = 12, height = 16)
```

### Selected trials

```{r fig.width = 12, fig.height = 3.5}
selected_trials = c(13, 7, 2, 23, 16, 29)

ggplot(data = data.exp1.long %>%
         filter(trial %in% selected_trials) %>%
         mutate(trial = factor(trial, levels = selected_trials),
                trial_label = as.integer(trial)),
       mapping = aes(x = condition,
                     y = response,
                     fill = condition)) +
  # mean judgments
  stat_summary(fun = "mean",
               geom = "bar",
               color = "black",
               alpha = 0.8,
               linewidth = 0.1) +
  # individual participant judgments
  geom_point(size = 1,
             stroke = 0.5,
             shape = 21,
             color = 'gray40',
             alpha = 0.5,
             position = position_jitter(height = 0, width = 0.2),
             show.legend = F) +
  # error bars
  stat_summary(fun.data = mean_cl_boot,
               geom = "linerange",
               color = "black",
               linewidth = 1) +
  # model predictions
  geom_point(data = data.exp1.model %>%
               filter(trial %in% selected_trials),
             mapping = aes(y = prediction),
             size = 3,
             stroke = 0.6,
             shape = 21,
             show.legend = F) +
  # trial images
  geom_custom(data = images.exp1 %>%
                filter(trial %in% selected_trials),
              inherit.aes = F,
              mapping = aes(data = grob, x = -Inf, y = Inf),
              grob_fun = function(x) rasterGrob(
                x, interpolate = T,
                vjust = 0, hjust = 0,
                height = unit(3.8, 'cm'))) +
  # trial image labels
  geom_text(aes(label = trial_label, x = 0.8, y = 126),
            color = 'black', size = 7,
            check_overlap = T) +
  facet_wrap(~ trial, nrow = 1, scales = 'free_x') +
  scale_fill_manual(values = exp1.condition_colors,
                    labels = exp1.condition_labels) +
  scale_y_continuous(name = 'Judgments',
                     breaks = seq(0, 100, 25),
                     expand = expansion(mult = 0.01)) +
  coord_cartesian(clip = "off",
                  ylim = c(0, 110)) +
  theme(text = element_text(size = 16),
        legend.position = 'bottom',
        legend.direction = 'horizontal',
        legend.text = element_text(
          size = 16,
          margin = margin(r = 1, l = 0.2, unit = 'cm')),
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        panel.grid.major.y = element_line(),
        plot.margin = margin(t = 3.9, l = 0, r = 0, b = 0, unit = 'cm'))

# ggsave(str_c(figures_dir, 'exp1_bar_select.pdf'),
#        width = 12, height = 3.5)
```

### Responsibility

```{r fig.width = 12, fig.height = 2.75}
scatterplot.exp1 = function(fit, x_label, y_label = '') {
  data.exp1.modeling %>%
    cbind(fit %>%
            fitted(newdata = data.exp1.modeling,
                   re_formula = NA) %>%
            as_tibble() %>%
            select(pred_resp_mean = Estimate)) %>%
    distinct(trial, pred_resp_mean) %>%
    left_join(data.exp1.resp %>%
                group_by(trial) %>%
                summarise(resp_mean = mean(response),
                          resp_low = smean.cl.boot(response)[2],
                          resp_high = smean.cl.boot(response)[3]),
              by = 'trial') %>%
    mutate(pred_resp_mean = pmax(pmin(pred_resp_mean, 100), 0),
           pred_resp_low = pred_resp_mean,
           pred_resp_high = pred_resp_mean) %>%
    scatterplot('pred_resp', x_label, 'resp', y_label)
}

plot.exp1.int = scatterplot.exp1(
  fit.exp1.int,
  '<span style="color:tan2">Intention</span>',
  'Responsibility')
plot.exp1.cf = scatterplot.exp1(
  fit.exp1.cf,
  '<span style="color:orchid3">Counterfactual</span>')
plot.exp1.cf_int = scatterplot.exp1(
  fit.exp1.cf_int,
  '<span style="color:tan2">Intention</span> + <br><span style="color:orchid3">Counterfactual</span>')
plot.exp1.cf_eff = scatterplot.exp1(
  fit.exp1.cf_eff,
  '<span style="color:aquamarine3">Effort</span> + <br><span style="color:orchid3">Counterfactual</span>')
plot.exp1.heuristic  = scatterplot.exp1(
  fit.exp1.heuristic,
  'Heuristic')

plot.exp1.int + plot.exp1.cf + plot.exp1.cf_int + plot.exp1.cf_eff + plot.exp1.heuristic +
  plot_layout(nrow = 1, widths = c(1, 1, 1, 1, 1)) &
  theme(plot.margin = margin(t = 0.2, r = 0.1, unit = 'cm'))

# ggsave(str_c(figures_dir, 'exp1_responsibility.pdf'),
#        width = 12, height = 2.75)
```

# Experiment 2

## Data

### Trial info

```{r}
trial.info.exp2 = read_csv(str_c(info_dir(2), 'model_predictions.csv')) %>%
  rename(cf_model = counterfactual,
         int_model = intention) %>%
  mutate(outcome = factor(outcome, levels = c('success', 'failure')),
         cf_model = cf_model * 100,
         int_model = int_model * 100,
         cf_model_recode = ifelse(outcome == 'success',
                                  100 - cf_model, cf_model),
         int_model_recode = ifelse(outcome == 'success',
                                   int_model, 100 - int_model))

images.exp2 = data.frame(trial = 1:nrow(trial.info.exp2)) %>%
  mutate(grob = map(.x = trial,
                    .f = ~ readPNG(str_c(figures_dir, 'trial_stills/experiment2/', .x, '.png'))))
```

### Responsibility

```{r}
data.exp2.resp = read_trials(2, 'responsibility')
data.exp2.resp.participants = read_participants(2, 'responsibility')

data.exp2.resp.participants %>% select(id, feedback) %>% datatable(rownames = F)
```

### Counterfactual

```{r}
data.exp2.cf = read_trials(2, 'counterfactual') %>%
  left_join(trial.info.exp2 %>%
              select(trial, outcome),
            by = 'trial') %>%
  mutate(cf_recode = ifelse(outcome == 'success',
                            100 - response, response))

data.exp2.cf.participants = read_participants(2, 'counterfactual')

data.exp2.cf.participants %>%
  select(id, feedback) %>%
  datatable(rownames = F)
```

### Intention

```{r}
data.exp2.int = read_trials(2, 'intention') %>%
  left_join(trial.info.exp2 %>%
              select(trial, outcome),
            by = 'trial') %>%
  mutate(int_recode = ifelse(outcome == 'success',
                             response, 100 - response))

data.exp2.int.participants = read_participants(2, 'intention')

data.exp2.int.participants %>%
  select(id, feedback) %>%
  datatable(rownames = F)
```

## Demographics

```{r}
print_demographics(data.exp2.resp.participants %>%
                     rbind(data.exp2.cf.participants) %>%
                     rbind(data.exp2.int.participants))
```

## Analysis

### Model components

```{r}
list(
  cf = list(judgment = data.exp2.cf %>%
              group_by(trial) %>%
              summarise(mean = mean(response)),
            model = trial.info.exp2 %>%
              select(trial, mean = cf_model)),
  int = list(judgment = data.exp2.int %>%
               group_by(trial) %>%
               summarise(mean = mean(response)),
             model = trial.info.exp2 %>%
               select(trial, mean = int_model))
) %>%
  map_dfr(.id = 'condition', function(x) {
    d = left_join(x$judgment, x$model,
                  by = 'trial',
                  suffix = c('_judgment', '_model'))
    tibble(r = cor(d$mean_judgment, d$mean_model),
           rmse = rmse(d$mean_judgment, d$mean_model))
  })
```

### Responsibility model

Setting up data:
```{r}
data.exp2.modeling = data.exp2.resp %>%
  left_join(data.exp2.cf %>%
              group_by(trial) %>%
              summarise(cf_recode = mean(cf_recode)),
            by = 'trial') %>%
  left_join(data.exp2.int %>%
              group_by(trial) %>%
              summarise(int_recode = mean(int_recode)),
            by = 'trial') %>%
  left_join(trial.info.exp2, by = 'trial')
```

Model fitting:
```{r}
fit.exp2.cf = brm(
  formula = response ~ 1 + cf_recode + (1 | id),
  data = data.exp2.modeling,
  iter = 2000,
  seed = 1,
  file = "cache/fit.exp2.cf")

summary(fit.exp2.cf)

fit.exp2.int = brm(
  formula = response ~ 1 + int_recode + (1 | id),
  data = data.exp2.modeling,
  iter = 2000,
  seed = 1,
  file = "cache/fit.exp2.int")

summary(fit.exp2.int)

fit.exp2.cf_int = brm(
  formula = response ~ 1 + cf_recode + int_recode + (1 | id),
  data = data.exp2.modeling,
  iter = 2000,
  seed = 1,
  file = "cache/fit.exp2.cf_int")

summary(fit.exp2.cf_int)

fit.exp2.heuristic = brm(
  formula = response ~ 1 + outcome + red_steps + blue_steps + box_steps + (1 | id),
  data = data.exp2.modeling,
  iter = 2000,
  seed = 1,
  file = "cache/fit.exp2.heuristic")

summary(fit.exp2.heuristic)
```

Model comparison:
```{r}
fit.exp2.cf = add_criterion(
  fit.exp2.cf,
  criterion = "loo",
  reloo = T,
  file = "cache/fit.exp2.cf")

fit.exp2.int = add_criterion(
  fit.exp2.int,
  criterion = "loo",
  reloo = T,
  file = "cache/fit.exp2.int")

fit.exp2.cf_int = add_criterion(
  fit.exp2.cf_int,
  criterion = "loo",
  reloo = T,
  file = "cache/fit.exp2.cf_int")

fit.exp2.heuristic = add_criterion(
  fit.exp2.heuristic,
  criterion = "loo",
  reloo = T,
  file = "cache/fit.exp2.heuristic")

loo_compare(fit.exp2.cf,
            fit.exp2.int,
            fit.exp2.cf_int,
            fit.exp2.heuristic)
```

Individual model fitting and comparison:
```{r eval = F}
fit.exp2.cf_individual = brm(
  formula = response ~ 1 + cf_recode,
  data = data.exp2.modeling %>% filter(id == 1),
  seed = 1,
  file = "cache/fit.exp2.cf_individual")

fit.exp2.int_individual = brm(
  formula = response ~ 1 + int_recode,
  data = data.exp2.modeling %>% filter(id == 1),
  seed = 1,
  file = "cache/fit.exp2.int_individual")

fit.exp2.cf_int_individual = brm(
  formula = response ~ 1 + cf_recode + int_recode,
  data = data.exp2.modeling %>% filter(id == 1),
  seed = 1,
  file = "cache/fit.exp2.cf_int_individual")

fit.exp2.heuristic_individual = brm(
  formula = response ~ 1 + outcome + red_steps + blue_steps + box_steps,
  data = data.exp2.modeling %>% filter(id == 1),
  seed = 1,
  file = "cache/fit.exp2.heuristic_individual")

data.exp2.fits = data.exp2.modeling %>%
  group_by(id) %>%
  nest() %>%
  ungroup() %>%
  mutate(
    fit_cf = map(.x = data,
                 .f = ~ update(fit.exp2.cf_individual,
                               newdata = .x,
                               seed = 1)),
    fit_int = map(.x = data,
                  .f = ~ update(fit.exp2.int_individual,
                                newdata = .x,
                                seed = 1)),
    fit_cf_int = map(.x = data,
                     .f = ~ update(fit.exp2.cf_int_individual,
                                   newdata = .x,
                                   seed = 1)),
    fit_heuristic = map(.x = data,
                        .f = ~ update(fit.exp2.heuristic_individual,
                                      newdata = .x,
                                      seed = 1))) %>%
  mutate(
    fit_cf = map(.x = fit_cf,
                 .f = ~ add_criterion(.x, criterion = "loo")),
    fit_int = map(.x = fit_int,
                  .f = ~ add_criterion(.x, criterion = "loo")),
    fit_cf_int = map(.x = fit_cf_int,
                     .f = ~ add_criterion(.x, criterion = "loo")),
    fit_heuristic = map(.x = fit_heuristic,
                        .f = ~ add_criterion(.x, criterion = "loo")),
    model_comparison = pmap(
      .l = list(cf = fit_cf,
                int = fit_int,
                cf_int = fit_cf_int,
                heuristic = fit_heuristic),
      .f = ~ loo_compare(..1, ..2, ..3, ..4)),
    best_model = map_chr(.x = model_comparison,
                         .f = ~ rownames(.) %>%
                           .[1]),
    best_model = factor(best_model,
                        levels = c("..1", "..2", "..3", "..4"),
                        labels = c("cf", "int", "cf_int", "heuristic")))

save(list = c('data.exp2.fits'),
     file = 'cache/exp2.model_fits.RData')
```

```{r}
load(file = 'cache/exp2.model_fits.RData')

data.exp2.fits %>%
  count(best_model) %>%
  arrange(desc(n))
```

### Measuring uncertainty

```{r}
dist_summary = bind_rows(
  data.exp1.cf %>% mutate(exp = 1, condition = 'cf',
                          dist = abs(response - 50)),
  data.exp1.int %>% mutate(exp = 1, condition = 'int',
                           dist = abs(response - 50)),
  data.exp2.cf %>% mutate(exp = 2, condition = 'cf',
                          dist = abs(response - 50)),
  data.exp2.int %>% mutate(exp = 2, condition = 'int',
                           dist = abs(response - 50))
)

dist_summary %>%
  group_by(exp, condition) %>%
  summarise(mean_dist = mean(dist),
            ci_low = mean_cl_boot(dist)$ymin,
            ci_high = mean_cl_boot(dist)$ymax) %>%
  mutate(exp = factor(exp))
```

Comparing intention judgments:
```{r}
fit.dist.int = brm(
  formula = dist ~ 1 + exp,
  data = dist_summary %>% filter(condition == 'int'),
  iter = 2000,
  seed = 1,
  file = "cache/fit.dist.int")

# summary(fit.dist.int)
hypothesis(fit.dist.int, "exp < 0", alpha = 0.05)
```

Comparing counterfactual judgments:
```{r}
fit.dist.cf = brm(
  formula = dist ~ 1 + exp,
  data = dist_summary %>% filter(condition == 'cf'),
  iter = 2000,
  seed = 1,
  file = "cache/fit.dist.cf")

# summary(fit.dist.cf)
hypothesis(fit.dist.cf, "exp < 0", alpha = 0.05)
```

## Figures

```{r}
exp2.conditions = c('int', 'cf', 'resp')

exp2.condition_labels = c('int' = 'Intention',
                          'cf' = 'Counterfactual',
                          'resp' = 'Responsibility')

exp2.condition_colors = c('int' = 'tan2',
                          'cf' = 'orchid3',
                          'resp' = 'deepskyblue')

data.exp2.long = data.exp2.resp %>%
  mutate(condition = 'resp') %>%
  rbind(data.exp2.cf %>%
          select(-outcome, -cf_recode) %>%
          mutate(condition = 'cf')) %>%
  rbind(data.exp2.int %>%
          select(-outcome, -int_recode) %>%
          mutate(condition = 'int')) %>%
  mutate(condition = factor(condition, levels = exp2.conditions))

data.exp2.model = trial.info.exp2 %>%
  select(trial, cf = cf_model, int = int_model) %>%
  left_join(data.exp2.modeling %>%
              cbind(fit.exp2.cf_int %>%
                      fitted(newdata = data.exp2.modeling,
                             re_formula = NA) %>%
                      as_tibble() %>%
                      select(pred = Estimate)) %>%
              distinct(trial, pred) %>%
              rename(resp = pred),
            by = 'trial') %>%
  pivot_longer(cols = -trial,
               names_to = 'condition',
               values_to = 'prediction') %>%
  mutate(condition = factor(condition, levels = exp2.conditions))
```

### All trials

```{r fig.width = 12, fig.height = 13}
ggplot(data = data.exp2.long,
       mapping = aes(x = condition,
                     y = response,
                     fill = condition)) +
  # mean judgments
  stat_summary(fun = "mean",
               geom = "bar",
               color = "black",
               alpha = 0.8,
               linewidth = 0.1) +
  # individual participant judgments
  geom_point(size = 1,
             stroke = 0.5,
             shape = 21,
             color = 'gray40',
             alpha = 0.5,
             position = position_jitter(height = 0, width = 0.2),
             show.legend = F) +
  # error bars
  stat_summary(fun.data = mean_cl_boot,
               geom = "linerange",
               color = "black",
               linewidth = 1) +
  # model predictions
  geom_point(data = data.exp2.model,
             mapping = aes(y = prediction),
             size = 3,
             stroke = 0.6,
             shape = 21,
             show.legend = F) +
  # trial images
  geom_custom(data = images.exp2,
              inherit.aes = F,
              mapping = aes(data = grob,
                            x = -Inf, y = Inf),
              grob_fun = function(x) rasterGrob(
                x, interpolate = T,
                vjust = 0, hjust = 0,
                height = unit(3.8, "cm"))) +
  # trial image labels
  geom_text(aes(label = trial,
                x = 0.68, y = 124),
            color = 'black',
            size = 6,
            check_overlap = T) +
  facet_wrap(~ trial, nrow = 4, scales = 'free_x') +
  scale_fill_manual(values = exp2.condition_colors,
                    labels = exp2.condition_labels) +
  scale_y_continuous(name = 'Judgments',
                     breaks = seq(0, 100, 25),
                     expand = expansion(mult = 0.01)) +
  coord_cartesian(clip = "off",
                  ylim = c(0, 110)) +
  theme(text = element_text(size = 16),
        legend.position = 'bottom',
        legend.direction = 'horizontal',
        legend.text = element_text(
          size = 16,
          margin = margin(l = 0.2, r = 1, unit = 'cm')),
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        panel.grid.major.y = element_line(),
        panel.spacing.y = unit(4.1, "cm"),
        plot.margin = margin(t = 3.9, r = 0, b = 0, l = 0, unit = 'cm'))

# ggsave(str_c(figures_dir, 'exp2_bar_all.pdf'),
#        width = 12, height = 13)
```

### Selected trials

```{r fig.width = 12, fig.height = 3.5}
selected_trials = c(6, 2, 16, 12, 23, 24)

ggplot(data = data.exp2.long %>%
         filter(trial %in% selected_trials) %>%
         mutate(trial = factor(trial, levels = selected_trials),
                trial_label = as.integer(trial)),
       mapping = aes(x = condition,
                     y = response,
                     fill = condition)) +
  # mean judgments
  stat_summary(fun = "mean",
               geom = "bar",
               color = "black",
               alpha = 0.8,
               linewidth = 0.1) +
  # individual participant judgments
  geom_point(size = 1,
             shape = 21,
             color = 'gray40',
             alpha = 0.5,
             position = position_jitter(height = 0, width = 0.2),
             show.legend = F) +
  # error bars
  stat_summary(fun.data = mean_cl_boot,
               geom = "linerange",
               color = "black",
               linewidth = 1) +
  # model predictions
  geom_point(data = data.exp2.model %>%
               filter(trial %in% selected_trials),
             mapping = aes(y = prediction),
             size = 3,
             stroke = 0.6,
             shape = 21,
             show.legend = F) +
  # trial images
  geom_custom(data = images.exp2 %>%
                filter(trial %in% selected_trials),
              inherit.aes = F,
              mapping = aes(data = grob,
                            x = -Inf, y = Inf),
              grob_fun = function(x) rasterGrob(
                x, interpolate = T,
                vjust = 0, hjust = 0,
                height = unit(3.7, 'cm'))) +
  # trial image labels
  geom_text(aes(label = trial_label,
                x = 0.7, y = 125),
            color = 'black',
            size = 7,
            check_overlap = T) +
  facet_wrap(~ trial, nrow = 1, scales = 'free_x') +
  scale_fill_manual(values = exp2.condition_colors,
                    labels = exp2.condition_labels) +
  scale_y_continuous(name = 'Judgments',
                     breaks = seq(0, 100, 25),
                     expand = expansion(mult = 0.01)) +
  coord_cartesian(clip = "off",
                  ylim = c(0, 110)) +
  theme(text = element_text(size = 16),
        legend.position = 'bottom',
        legend.direction = 'horizontal',
        legend.text = element_text(
          size = 16,
          margin = margin(l = 0.2, r = 1, unit = 'cm')),
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        panel.grid.major.y = element_line(),
        plot.margin = margin(t = 3.8, l = 0,
                             r = 0, b = 0, unit = 'cm'))

# ggsave(str_c(figures_dir, 'exp2_bar_select.pdf'),
#        width = 12, height = 3.5)
```

### Responsibility

```{r fig.width = 12, fig.height = 3}
resp_scatterplot2 = function(fit, x_label, y_label = '') {
  data.exp2.modeling %>%
    cbind(fit %>%
            fitted(newdata = data.exp2.modeling,
                   re_formula = NA) %>%
            as_tibble() %>%
            select(pred_resp_mean = Estimate)) %>%
    distinct(trial, pred_resp_mean) %>%
    left_join(data.exp2.resp %>%
                group_by(trial) %>%
                summarise(resp_mean = mean(response),
                          resp_low = smean.cl.boot(response)[2],
                          resp_high = smean.cl.boot(response)[3]),
              by = 'trial') %>%
    mutate(pred_resp_mean = pmax(pmin(pred_resp_mean, 100), 0),
           pred_resp_low = pred_resp_mean,
           pred_resp_high = pred_resp_mean) %>%
    scatterplot('pred_resp', x_label, 'resp', y_label)
}

plot.exp2.int = resp_scatterplot2(
  fit.exp2.int,
  '<span style="color:tan2">Intention</span>',
  'Responsibility')
plot.exp2.cf = resp_scatterplot2(
  fit.exp2.cf,
  '<span style="color:orchid3">Counterfactual</span>')
plot.exp2.cf_int = resp_scatterplot2(
  fit.exp2.cf_int,
  '<span style="color:tan2">Intention</span> + <span style="color:orchid3">Counterfactual</span>')
plot.exp2.heuristic = resp_scatterplot2(
  fit.exp2.heuristic,
  'Heuristic')

plot.exp2.int + plot.exp2.cf + plot.exp2.cf_int + plot.exp2.heuristic +
  plot_layout(nrow = 1) &
  theme(plot.margin = margin(0))

# ggsave(str_c(figures_dir, 'exp2_responsibility.pdf'),
#        width = 12, height = 3)
```

# Experiment 3

## Data

### Trial info

```{r}
trial.info.exp3 = read_csv(str_c(info_dir(3), 'model_predictions.csv')) %>%
  select(-intention_t1) %>%
  rename(cf_belief_no_prior = counterfactual_belief_no_prior,
         cf_belief_with_prior = counterfactual_belief_with_prior,
         cf = counterfactual,
         int_no_prior = intention_t2_no_prior,
         int_with_prior = intention_t2_with_prior) %>%
  mutate(trial_type = factor(trial_type,
                             levels = c('helping', 'hindering')),
         across(c(-trial, -trial_type), ~ .x * 100))

images.exp3.t1 = data.frame(trial = 1:nrow(trial.info.exp3)) %>%
  mutate(grob = map(.x = trial,
                    .f = ~ readPNG(str_c(figures_dir, 'trial_stills/experiment3/', .x, '-1.png'))))

images.exp3.t2 = data.frame(trial = 1:nrow(trial.info.exp3)) %>%
  mutate(grob = map(.x = trial,
                    .f = ~ readPNG(str_c(figures_dir, 'trial_stills/experiment3/', .x, '-2.png'))))
```

### Trials

```{r}
data.exp3 = read_trials(3, 'no_prior') %>%
  mutate(condition = 'no_prior') %>%
  rbind(read_trials(3, 'with_prior') %>%
          select(-judge_t1) %>%
          mutate(condition = 'with_prior')) %>%
  mutate(condition = factor(condition,
                            levels = c('no_prior', 'with_prior'))) %>%
  left_join(trial.info.exp3 %>%
              select(trial, trial_type))

contrasts(data.exp3$condition) = c(-1, 1)
contrasts(data.exp3$trial_type) = c(-1, 1)
```

### Participants

```{r}
data.exp3.participants = read_participants(3, 'no_prior') %>%
  mutate(condition = 'no_prior') %>%
  rbind(read_participants(3, 'with_prior') %>%
          mutate(condition = 'with_prior'))

data.exp3.participants %>%
  select(id, feedback) %>%
  datatable(rownames = F)
```

## Demographics

```{r}
print_demographics(data.exp3.participants)
```

## Analysis

### Prior expectations

```{r}
data.exp3.prior_judge = read_trials(3, 'with_prior') %>%
  select(id, trial, judge = judge_t1) %>%
  left_join(trial.info.exp3 %>% select(trial, trial_type))

data.exp3.prior_judge %>%
  group_by(trial_type) %>%
  summarise(mean = mean(judge), sd = sd(judge))

t.test(formula = judge ~ trial_type, data = data.exp3.prior_judge)
```

### Responsibility data

```{r}
data.exp3 %>%
  pivot_longer(cols = c(resp_red, resp_blue), names_to = "agent", values_to = "resp") %>%
  group_by(condition, agent) %>%
  reframe({
    m = mean_cl_boot(resp, conf.int = 0.95)
    tibble(mean = round(m$y, 1), lower = round(m$ymin, 1), upper = round(m$ymax, 1))
  })

data.exp3 %>%
  pivot_longer(cols = c(resp_red, resp_blue), names_to = "agent", values_to = "resp") %>%
  group_by(trial_type, agent) %>%
  reframe({
    m = mean_cl_boot(resp, conf.int = 0.95)
    tibble(mean = round(m$y, 1), lower = round(m$ymin, 1), upper = round(m$ymax, 1))
  })
```

### Responsibility model

Setting up data:
```{r}
data.exp3.modeling = data.exp3 %>%
  select(-trial_type) %>%
  left_join(trial.info.exp3 %>%
              select(trial, starts_with('cf')) %>%
              pivot_longer(cols = starts_with('cf_belief'),
                           names_to = 'condition',
                           names_prefix = 'cf_belief_',
                           values_to = 'cf_belief') %>%
              mutate(condition = factor(condition,
                                        levels = c('no_prior', 'with_prior'))),
            by = c('trial', 'condition')) %>%
  mutate(int = 100 - int)
```

Model fitting:
```{r}
fit.exp3.blue = brm(
  formula = resp_blue ~ 1 + cf + int + (1 | id),
  data = data.exp3.modeling,
  iter = 2000,
  seed = 1,
  file = "cache/fit.exp3.blue")

summary(fit.exp3.blue)

fit.exp3.red = brm(
  formula = resp_red ~ 1 + cf_belief + (1 | id),
  data = data.exp3.modeling,
  iter = 2000,
  seed = 1,
  file = "cache/fit.exp3.red")

summary(fit.exp3.red)
```

## Figures

```{r}
exp3.judgment_labels = c('int' = 'Intention',
                         'resp_blue' = 'Responsibility (Blue)',
                         'resp_red' = 'Responsibility (Red)')

exp3.judgment_names = names(exp3.judgment_labels)

exp3.judgment_colors_x = c('int.no_prior' = '#FAE1D0',
                           'int.with_prior' = 'tan2',
                           'resp_blue.no_prior' = '#C7EEFE',
                           'resp_blue.with_prior' = 'deepskyblue',
                           'resp_red.no_prior' = '#FDCBCD',
                           'resp_red.with_prior' = 'brown1')
exp3.judgment_colors = c('int' = 'tan2',
                         'resp_blue' = 'deepskyblue',
                         'resp_red' = 'brown1')

exp3.condition_alphas = c(0.3, 0.8)
exp3.condition_labels = c('no_prior' = 'No prior',
                          'with_prior' = 'With prior')
exp3.trial_type_labels = c('helping' = 'Helping trials',
                           'hindering' = 'Hindering trials')

dodge_width = 0.9

trial_labels.exp3 = data.frame(
  trial = rep(1:nrow(trial.info.exp3), each = 2),
  trial_type = rep(c('hindering', 'helping'),
                   each = nrow(trial.info.exp3)),
  t = rep(1:2, times = nrow(trial.info.exp3)),
  tt = rep(c('A', 'B'), times = nrow(trial.info.exp3))) %>%
  mutate(trial_label = str_c(trial, tt))
```

```{r}
data.exp3.long = data.exp3 %>%
  pivot_longer(cols = c(starts_with('int'), starts_with('resp')),
               names_to = 'judgment_type',
               values_to = 'response') %>%
  mutate(judgment_type = factor(judgment_type,
                                levels = exp3.judgment_names),
         condition = factor(condition,
                            levels = c('no_prior', 'with_prior')))

data.exp3.model = trial.info.exp3 %>%
  select(trial, trial_type, starts_with('int_')) %>%
  pivot_longer(cols = c(-trial, -trial_type),
               names_to = c(".value", "condition"),
               names_pattern = "^(int)_(.*)$") %>%
  left_join(data.exp3.modeling %>%
              cbind(fit.exp3.red %>%
                      fitted(newdata = data.exp3.modeling,
                             re_formula = NA) %>%
                      as_tibble() %>%
                      select(pred = Estimate)) %>%
              distinct(trial, condition, pred) %>%
              rename(resp_red = pred),
            by = c('trial', 'condition')) %>%
  left_join(data.exp3.modeling %>%
              cbind(fit.exp3.blue %>%
                      fitted(newdata = data.exp3.modeling,
                             re_formula = NA) %>%
                      as_tibble() %>%
                      select(pred = Estimate)) %>%
              group_by(trial, condition) %>%
              summarise(resp_blue = mean(pred)) %>%
              ungroup(),
            by = c('trial', 'condition')) %>%
  pivot_longer(cols = c(-trial, -condition, -trial_type),
               names_to = 'judgment_type',
               values_to = 'response')

data.exp3.means = data.exp3 %>%
  group_by(trial, condition) %>%
  summarise(resp_red_mean = mean(resp_red),
            resp_red_low = smean.cl.boot(resp_red)[2],
            resp_red_high = smean.cl.boot(resp_red)[3],
            resp_blue_mean = mean(resp_blue),
            resp_blue_low = smean.cl.boot(resp_blue)[2],
            resp_blue_high = smean.cl.boot(resp_blue)[3]) %>%
  ungroup()
```

### Aggregated judgments

```{r fig.width = 12, fig.height = 5}
selected_trials = c(12, 1)

ggplot(data = data.exp3.long,
       mapping = aes(x = judgment_type,
                     y = response,
                     group = condition)) +
  # mean judgments
  stat_summary(mapping = aes(fill = judgment_type,
                             alpha = condition),
               fun = "mean",
               geom = "bar",
               color = "black",
               linewidth = 0.2,
               position = position_dodge2(width = dodge_width,
                                          padding = 0.1)) +
  scale_alpha_manual(values = exp3.condition_alphas,
                     labels = exp3.condition_labels) +
  scale_fill_manual(values = exp3.judgment_colors,
                    labels = exp3.judgment_labels) +
  new_scale_fill() +
  # individual participant judgments
  geom_point(mapping = aes(fill = interaction(judgment_type, condition)),
             size = 1,
             shape = 21,
             color = 'gray50',
             alpha = 0.2,
             position = position_jitterdodge(dodge.width = dodge_width),
             show.legend = F) +
  # error bars
  stat_summary(fun.data = mean_cl_boot,
               geom = "linerange",
               color = "black",
               alpha = 0.8,
               linewidth = 1,
               position = position_dodge(width = dodge_width),
               show.legend = F) +
  # model predictions
  geom_point(data = data.exp3.model %>%
               group_by(trial_type, condition, judgment_type) %>%
               summarise(response = mean(response)),
             mapping = aes(fill = interaction(judgment_type, condition)),
             size = 3,
             stroke = 0.6,
             shape = 21,
             position = position_dodge(width = dodge_width),
             show.legend = F) +
  scale_fill_manual(values = exp3.judgment_colors_x) +
  # example trial images A
  geom_custom(data = images.exp3.t1 %>%
                filter(trial %in% selected_trials) %>%
                merge(trial.info.exp3 %>%
                        select(trial, trial_type)),
              inherit.aes = F,
              mapping = aes(data = grob, x = -Inf, y = Inf),
              grob_fun = function(x) rasterGrob(
                x, interpolate = T,
                vjust = 0, hjust = -0.3,
                height = unit(4.5, 'cm'))) +
  # example trial images B
  geom_custom(data = images.exp3.t2 %>%
                filter(trial %in% selected_trials) %>%
                merge(trial.info.exp3 %>%
                        select(trial, trial_type)),
              inherit.aes = F,
              mapping = aes(data = grob, x = -Inf, y = Inf),
              grob_fun = function(x) rasterGrob(
                x, interpolate = T,
                vjust = 0, hjust = -1.35,
                height = unit(4.5, 'cm'))) +
  # example trial image labels
  geom_text(data = data.frame(trial_type = 'helping',
                              label = 'Example'),
            inherit.aes = F,
            mapping = aes(label = label,
                          x = 0.1, y = 160),
            angle = 90,
            size = 6, 
            show.legend = F) +
  # trial type labels
  geom_rect(data = data.frame(xmin = -Inf, xmax = Inf,
                              ymin = 207, ymax = 230),
            mapping = aes(xmin = xmin, xmax = xmax,
                          ymin = ymin, ymax = ymax),
            inherit.aes = F, fill = 'gray90') +
  geom_text(data = data.frame(trial = rep(selected_trials, each = 2),
                              t = rep(1:2, times = 2),
                              label = rep(c('A', 'B'), times = 2)),
            aes(label = label, x = (t-1)*1.28 + 0.87, y = 121),
            inherit.aes = F,
            color = 'black',
            size = 8,
            show.legend = F) +
  facet_wrap(~ trial_type,
             labeller = as_labeller(exp3.trial_type_labels)) +
  scale_y_continuous(name = 'Judgments',
                     breaks = seq(0, 100, 25),
                     expand = expansion(mult = 0.01)) +
  coord_cartesian(clip = "off",
                  ylim = c(0, 110),
                  xlim = c(1, 3)) +
  theme(text = element_text(size = 16),
        legend.position = 'bottom',
        legend.direction = 'horizontal',
        legend.text = element_text(
          size = 16,
          margin = margin(l = 0.2, r = 0.4, unit = 'cm')),
        legend.margin = margin(l = 0.5, r = 0.5, unit = 'cm'),
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(
          size = 16, face = 'bold',
          margin = margin(b = 5.1, t = 0.3, unit = 'cm')),
        panel.grid.major.y = element_line(),
        plot.margin = margin(0))

# ggsave(str_c(figures_dir, 'exp3_bar_aggregate.pdf'),
#        width = 12, height = 5)
```

### All trials

```{r fig.width = 12, fig.height = 13}
ggplot(data = data.exp3.long,
       mapping = aes(x = judgment_type,
                     y = response,
                     group = condition)) +
  stat_summary(mapping = aes(fill = judgment_type,
                             alpha = condition),
               fun = "mean",
               geom = "bar",
               color = "black",
               linewidth = 0.1,
               position = position_dodge(width = dodge_width)) +
  scale_alpha_manual(values = exp3.condition_alphas,
                     labels = exp3.condition_labels) +
  scale_fill_manual(values = exp3.judgment_colors,
                    labels = exp3.judgment_labels,
                    na.translate = F) +
  new_scale_fill() +
  geom_point(mapping = aes(fill = interaction(judgment_type, condition)),
             size = 1,
             shape = 21,
             color = "gray40",
             alpha = 0.5,
             position = position_jitterdodge(dodge.width = dodge_width),
             show.legend = F) +
  stat_summary(fun.data = mean_cl_boot,
               geom = "linerange",
               color = "black",
               alpha = 0.8,
               linewidth = 1,
               position = position_dodge(width = dodge_width),
               show.legend = F) +
  geom_point(data = data.exp3.model,
             mapping = aes(fill = interaction(judgment_type, condition)),
             size = 3,
             stroke = 0.6,
             shape = 21,
             position = position_dodge(width = dodge_width),
             show.legend = F) +
  scale_fill_manual(values = exp3.judgment_colors_x) +
  geom_custom(data = images.exp3.t1,
              mapping = aes(data = grob, x = -Inf, y = Inf),
              inherit.aes = F,
              grob_fun = function(x) rasterGrob(
                x, interpolate = T,
                vjust = 0, hjust = -0.1,
                height = unit(3.6, 'cm'))) +
  geom_custom(data = images.exp3.t2,
              mapping = aes(data = grob, x = -Inf, y = Inf),
              inherit.aes = F,
              grob_fun = function(x) rasterGrob(
                x, interpolate = T,
                vjust = 0, hjust = -1.12,
                height = unit(3.6, 'cm'))) +
  geom_text(data = trial_labels.exp3,
            aes(label = trial_label,
                x = (t-1)*1.5 + 0.73,
                y = 122),
            inherit.aes = F,
            color = 'black',
            size = 6,
            check_overlap = T,
            show.legend = F) +
  facet_wrap(~ trial, nrow = 4, scales = "free_x") +
  scale_y_continuous(name = 'Judgments',
                     breaks = seq(0, 100, 25),
                     expand = expansion(mult = 0.01)) +
  coord_cartesian(clip = "off", ylim = c(0, 110)) +
  theme(text = element_text(size = 16),
        legend.position = 'bottom',
        legend.direction = 'horizontal',
        legend.text = element_text(
          size = 16,
          margin = margin(l = 0.2, r = 0.4, unit = 'cm')),
        legend.margin = margin(l = 0.5, r = 0.5, unit = 'cm'),
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        panel.grid.major.y = element_line(),
        panel.spacing.y = unit(3.9, "cm"),
        plot.margin = margin(t = 3.7, l = 0, r = 0, b = 0, unit = 'cm'))

# ggsave(str_c(figures_dir, 'exp3_bar_all.pdf'),
#        width = 12, height = 13)
```

### Responsibility

Red:
```{r fig.width = 9, fig.height = 4}
data.exp3.red = data.exp3.means %>%
  left_join(data.exp3.model %>%
              filter(judgment_type == 'resp_red') %>%
              rename(pred_mean = response),
            by = c('trial', 'condition'))

r = cor(data.exp3.red$pred_mean, data.exp3.red$resp_red_mean)
rmse = rmse(data.exp3.red$pred_mean, data.exp3.red$resp_red_mean)

plot.exp3.red = ggplot(data = data.exp3.red,
                       aes(x = pred_mean,
                           y = resp_red_mean,
                           fill = condition)) +
  geom_abline(intercept = 0, slope = 1,
              linetype = 2, size = 0.2) +
  geom_linerange(size = 0.7,
                 mapping = aes(ymin = resp_red_low,
                               ymax = resp_red_high),
                 color = 'lightgray') +
  geom_point(shape = 21, size = 3, stroke = 0.1) +
  geom_text(label = sprintf('RMSE = %.2f\nr = %.2f', rmse, r),
            hjust = 0, x = 0, y = 95,
            size = 4.5, check_overlap = T) +
  scale_x_continuous(name = 'Counterfactual Belief Model',
                     limits = c(0, 100)) +
  scale_y_continuous(name = 'Responsibility for Red',
                     limits = c(0, 100)) +
  scale_fill_manual(values = c('no_prior' = exp3.judgment_colors_x[['resp_red.no_prior']],
                               'with_prior' = exp3.judgment_colors_x[['resp_red.with_prior']]),
                    labels = c('No prior', 'With prior')) +
  theme(legend.position = c(0.9, 0.2),
        legend.title = element_blank(),
        text = element_text(size = 16),
        axis.title.x = element_markdown(size = 16),
        axis.title.y = element_markdown(size = 16),
        aspect.ratio = 1)
```

Blue:
```{r}
data.exp3.blue = data.exp3.means %>%
  left_join(data.exp3.model %>%
              filter(judgment_type == 'resp_blue') %>%
              rename(pred_mean = response),
            by = c('trial', 'condition'))

r = cor(data.exp3.blue$pred_mean, data.exp3.blue$resp_blue_mean)
rmse = rmse(data.exp3.blue$pred_mean, data.exp3.blue$resp_blue_mean)

plot.exp3.blue = ggplot(data = data.exp3.blue,
                        aes(x = pred_mean,
                            y = resp_blue_mean,
                            fill = condition)) +
  geom_abline(intercept = 0, slope = 1,
              linetype = 2, size = 0.2) +
  geom_linerange(size = 0.7,
                 mapping = aes(ymin = resp_blue_low,
                               ymax = resp_blue_high),
                 color = 'lightgray') +
  geom_point(shape = 21, size = 3, stroke = 0.1) +
  geom_text(label = sprintf('RMSE = %.2f\nr = %.2f', rmse, r),
            hjust = 0, x = 0, y = 95,
            size = 4.5, check_overlap = T) +
  scale_x_continuous(name = 'Intention + Counterfactual Model',
                     limits = c(0, 100)) +
  scale_y_continuous(name = 'Responsibility for Blue',
                     limits = c(0, 100)) +
  scale_fill_manual(values = c('no_prior' = exp3.judgment_colors_x[['resp_blue.no_prior']],
                               'with_prior' = exp3.judgment_colors_x[['resp_blue.with_prior']]),
                    labels = c('No prior', 'With prior')) +
  theme(legend.position = c(0.9, 0.2),
        legend.title = element_blank(),
        legend.text = element_text(size = 14),
        text = element_text(size = 16),
        axis.title.x = element_markdown(size = 16),
        axis.title.y = element_markdown(size = 16),
        aspect.ratio = 1)
```

```{r fig.width = 9, fig.height = 4}
plot.exp3.blue + plot_spacer() + plot.exp3.red +
  plot_layout(widths = c(8, 1, 8)) &
  theme(plot.margin = margin(t = 0, r = 0.3, b = 0, l = 0, unit = 'cm'))

# ggsave(str_c(figures_dir, 'exp3_responsibility.pdf'),
#        width = 9, height = 4)
```

# Session info
```{r}
sessionInfo()
```
